Post-Disturbance Tree Community Trajectories in a Neotropical Forest

### 2018-06-05 WholePlotTrajectories.Rmd
relecture

  To disentangle the ecological processes driving forests dynamics we examined the the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
  Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.

  We highlighted a decoupling between functional and taxonomic trajectories, translating a fast recovery of communities functioning while their taxonomic recovery was slowed by the recruitment of infrequent species specific to mature forests.
  The IDH consistently predicted communities functional trajectories but poorly reflected the taxonomic ones, blurried by changes in the functional redundancy entailing competitive exclusion.
  Although communities consistently recovered their functioning after 30 years, initially infrequent species proved long to recover which impacted the taxonomic structure and functoinal redundancy over several decades.
  Our results acknowledged the need of decades-long recovery cycles to restore the resilience of the community and prevent species loss in the long term.

@@

Both communities taxonomic and functional diversity and composition proved resilient, following similar humped shaped trajectories returning towards initial values.
The resilience of communities functional structure, that is the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003].
It besides suggested that recovery trajectories depend on the initial composition eventually restored that would the the asymptot of the humped shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].

Although both communities taxonomic and functional characteristics proved resilient and followed similar humped shaped trajectories, the taxonomic recovery systematically lagged behind the functional one.
The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness were enhanced after disturbance, and which rapidly recovered their initial functional structure.
However, communities evenness remained high so infrequent species still missed to the taxonomic recovery that proved all the more longer to recover that unrecovered species would be functionally redundant and probably underwent competition processes.

@@

Validating the IDH, communities trajectories confirmed the diversity increase after disturbance through an enhanced growth of otherwise less favored species.
This was strong regarding the functional diversity but more blurry regarding their taxonomic structure.

The taxonomic richness was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a], while it substantially increased after low intensity disturbance.
Disturbance also enhanced taxonomic evenness, but the disturbance intensity was only weakly and lately correlated to communities eveness (*i.e.* Simpson diversity).
Contrastingly, disturbance intensity consistently predicted the significant increase of communities functional diversity for 30 years.
Disturbance entailed specific turnover within communities, either among pre-disturbance survivors or among newly recruited trees.
As the composition of old-growth survivors proved to mirror initial communities [@Herault2018], disturbance likely impacted trees recruitment through the enhanced growth and survival of previously infrequent species and functional types.
Consistently, disturbance resulted in increasing taxonomic dissimilarity compared to pre-disturbance communities and significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
Disturbance then caused a reorganization of the typical high dominance structure of hyperdiverse mature forests after disturbance, benefiting to pioneers and light demanding species. 
Likely, the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but are excluded in mature forests through the cmpetition with long-lived, more resistant and shade bearers.
Therefore, consistently with the IDH the trajectories after disturbance relied on the environmental niches made available by disturbance filled by species from a restrictied functional range.
Recruited species therefore differed from pre-disturbance ones, constituting a community all the more diversified that the disturbance was intense [@Molino2001].

@@

Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity made central the role of functional redundancy.
Functional redundancy, determinant of communities resilience [@Trenbath1999; @Elmqvist2003; @Diaz2005], seemed not to have recovered 30 years after disturbance.

Communities with high functional redundancy are expected to be able to lose species with a limited impact on their functional structure [@Carmona2016].
Functional redundancy therefore mitigates the impacts of disturbance, but we expect the redundancy itself to be reduced or re-organized within the functional space.
Although the overall functional redundancy did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
This restricted redundancy decreased according to the disturbance intensity before returning towards the initial value but had not recovered for any disturbed community after 30 years.
The recovery of the redundancy within the initial functional space explained the hampered recovery of infrequent species specific to old-growth forests that would be functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
The unachieved humped shaped redundancy trajectory meant a decrease of the resilience of pre-disturbance communities and an increase in the resilience of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
After disturbance the chances were higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013], which highly questioned the completeness of the resilience [@Chazdon2003a].
Specificaly, species contingent to undisturbed forests might be impaired, threatening their maintenance, the risk to loose cornerstone species and trigger unexpected ecological consequences would be higher [@Jones1994;@Diaz2005;@Gardner2007].

@@

our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbance.
These trajectories shed light on their rapid functional recovery but slower and more variable taxonomic dynamics.
Consistently with the IDH, functional trajectories were constrained by determined processes favoring pioneers and light-demanding species after disturbance.
The taxonomic dynamics however proved more stochastic due to the recovery of infrequent, functionally redundant species hampered by limiting similarity processes.
Although communities functioning rapidly recovered, their taxonomic structure and functional redundancy remained altered for several decades.
The trajectories therefore demonstrated a decades-long resilience of tropical forests in the face of quite intense disturbance [@Gourlet-Fleury2005].
Within the mechanisms of this resilience the recruitment processes proved central, in accordance with the IDH, and would deserved closer focus to clarify their determinisms and ecological rules.
The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.

>>>>

    To disentangle the ecological processes driving forests dynamics we examined the the trajectories over 30 years after a gradient of logging and thinning disturbance, focusing on communities taxonomic and functional composition and diversity.
    Specifically we analysed the trajectories of communities taxonomic richness and evennes and functional composition, diversity and redundancy based on 7 key functional traits assessing species ecology.
    We highlighted a decoupling between functional and taxonomic trajectories, translating a fast recovery of communities functioning while their taxonomic recovery was slowed by the recruitment of infrequent species specific to mature forests.
    The IDH consistently predicted communities functional trajectories but poorly reflected the taxonomic ones, blurried by changes in the functional redundancy entailing competitive exclusion.
    Although communities consistently recovered their functioning after 30 years, initially infrequent species proved long to recover which impacted the taxonomic structure and functoinal redundancy over several decades.
    Our results acknowledged the need of decades-long recovery cycles to restore the resilience of the community and prevent species loss in the long term.

@@

Both communities taxonomic and functional diversity and composition proved resilient, following similar humped shaped trajectories returning towards initial values.
The resilience of communities functional structure, that is the most direct link between biodiversity and ecosystem functioning [@Diaz2005], meant a consistent recovery of ecosystem processes in the long term [@Guariguata2001].
The resilience of communities taxonomy meant the maintenance of their initial differences in composition and diversity, suggesting the existence of multiple stable equilibria as assumed for highly diverse and productive ecosystems [@Chase2003], and the dependency of recovery trajectories on the initial composition to be recovered that would the the asymptot of the humped shaped dynamics [@Hubbell1999; @Molino2001; @Anderson2007; @Baraloto2012a].

Although both communities taxonomic and functional characteristics proved resilient and followed similar humped shaped trajectories, the taxonomic recovery systematically lagged behind the functional one.
The decoupling between functional and taxonomic dynamics has already been observed for grasslands [@Tilman1997; @Mouillot2011] and more recently for tropical forests [@Lohbeck2015; @Guariguata2001].
According to the "vegetation quantity effect" [@Grime1998] functional trajectories rely on the pool of dominant species, which diversity and evenness increased after disturbance but rapidly recovered their initial functional structure.
Although the pool of dominant species recovered quickly, communities evenness remained high so infrequent species still missed to the taxonomic recovery.
Taxonomic recovery was then all the more longer that unrecovered species would be functionally redundant and probably underwent selective competition processes.

@@

In accordance with the IDH the trajectories confirmed the increase of communities diversity after disturbance with the favored growth of otherwise less favored species.
However, although it was clear for the functional diversity this tendency was more blurry regarding communities taxonomic structure.

The taxonomic richness was enhanced after low disturbance but it was weakly or negatively impacted by intense disturbance, as observed on several post logging surveys [@Cannon1998; @Baraloto2012a].
The taxonomic evenness also increased after disturbance, but the correlation between the disturbance intensity and the diversity increase remained weak and valid only after 20 years (*i.e.* Simpson diversity).

Contrastingly the functional diversity increase after disturbance was determined by the disturbance intensity all along the 30 years monitored.
We already demonstrated that disturbance primarily entailed turnover of specific within communities, either among pre-disturbance survivors or among newly recruited trees.
It wasdemonstrated that the composition of old-growth survivors proved to mirror initial communities [@Herault2018], so disturbance rather impacted trees recruitment with the enhanced growth and survival of previously infrequent species and functional types.
Consistently disturbance resulted in an increase of the taxonomic dissimilarity between communities and their pre-disturbance composition, and in significant functional shifts towards resource-acquisitive strategies (sharp increase in the SLA, leaf thickness and bark thickness and decrease in wood density, leaf toughness and maximum height) [@Westoby1998; @Wright2004; @Reich2014].
Disturbance entailed a reorganization of the typical high dominance structure of hyperdiverse mature forests benefiting to pioneers and light demanding species: the changes in abiotic environment and competitive pressure favored pioneers which outcompete other species in non limiting resources but were excluded in mature forests by long-lived, more resistant and shade tolerant species.
Corroborating the IDH, the trajectories after disturbance relied on the environmental niches made available and filled by species from a restrictied functional range.
Recruited species therefore differed from pre-disturbance ones, constituting a community all the more diverse that the disturbance was intense [@Molino2001].

@@

Both the lag between taxonomic and functional recovery and the middling consistency of the IDH regarding communities taxonomic diversity were explained by the incomplete recovery of communities functional redundancy, which besides implies that communities resilience remained altered 30 years after disturbance [@Trenbath1999; @Elmqvist2003; @Diaz2005].

Functional redundancy mitigates the loss of species on communities functional structure [@Carmona2016], and therefore reduces the impact on disturbance.
The functional redundancy itself however might be reduced or re-organized within the functional space  and alter communities resilience.
Although the functional redundancy considered in the whole commnities functional space did not follow consistent trajectories after disturbance, the redundancy restricted to the functional space of the initial community clearly followed humped shaped trajectories.
This restricted redundancy first decreased, all the more so that the disturbance was intense, and then returned towards the initial value but this had not recovered for any disturbed community after 30 years.
After disturbance communities then differently occupied the functional space, as it was suggested by the shifts observed for the functional diversity and traits trajectories, before recovering the specific highly redundant initial functional space.
This increasing redundancy along the trajectory explained the hampered recovery of infrequent species specific to old-growth forests, as they were likely functionally redundant and therefore underwent competitive pressures limiting species functional similarity [@Mayfield2010].
The unachieved recovery of the functional redundancy meant a decreased resilience of pre-disturbance communities and an increased of disturbance-specific communities (corresponding to lower richness and higher dominance of pioneer species).
After disturbance the chances were therefore higher to have long lasting or self-maintained compositional changes in favor of disturbance resistant species, lianas or epiphytes [@Haddad2008; @Burslem2000; @Martin2013].
The recovery was not complete until the functional redundancy and resilience of the initial communities remained altered.
Specificaly, such incomplete recovery would threaten the maintenance of species contingent to undisturbed forests and increase the risk of cornerstone species loss and unexpected ecological consequences [@Jones1994;@Chazdon2003a;@Diaz2005;@Gardner2007].

@@

Our study defined the decoupled functional and taxonomic trajectories of tropical rainforests after disturbancethat demonstrated a rapid functional recovery but a slower and more variable taxonomic one.
Consistently with the IDH, functional trajectories were driven by deterministic processes favoring pioneers and light-demanding species after disturbance.
The following  functional shifts entailed a re-organization of communities functional redundancy in the functional space that proved long to recover and explained the more variable and longer taxonomic trajectories as shade tolerant species of old-growth forests faced functional redundancy and limiting similarity processes.
The resilience of tropical forests then proved long in the face of quite intense disturbance and did not preclude the settling of a persistent disturbanc-specific community [@Gourlet-Fleury2005].
Within the mechanisms of this resilience the recruitment processes proved central, in accordance with the IDH, and would deserve closer focus to specify their ecological drivers.
The disturbance range however stayed within the spectrum of selective logging, with a forest cover remaining all along the experiment, and the response mechanisms would probably be much different after harder disturbance.

### 2018-06-05 WholePlotTrajectories.tex
relecture

Both communities taxonomic and functional diversity and composition
proved resilient, following similar humped shaped trajectories returning
towards initial values. The resilience of communities functional
structure, that is the most direct link between biodiversity and
ecosystem functioning \citep{Diaz2005}, meant a consistent recovery of
ecosystem processes in the long term \citep{Guariguata2001}. The
resilience of communities taxonomy meant the maintenance of their
initial differences in composition and diversity, suggesting the
existence of multiple stable equilibria as assumed for highly diverse
and productive ecosystems \citep{Chase2003}. It besides suggested that
recovery trajectories depend on the initial composition eventually
restored that would the the asymptot of the humped shaped dynamics
\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.

Although both communities taxonomic and functional characteristics
proved resilient and followed similar humped shaped trajectories, the
taxonomic recovery systematically lagged behind the functional one. The
decoupling between functional and taxonomic dynamics has already been
observed for grasslands \citep{Tilman1997, Mouillot2011} and more
recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
According to the ``vegetation quantity effect'' \citep{Grime1998}
functional trajectories rely on the pool of dominant species, which
diversity and evenness were enhanced after disturbance, and which
rapidly recovered their initial functional structure. However,
communities evenness remained high so infrequent species still missed to
the taxonomic recovery that proved all the more longer to recover that
unrecovered species would be functionally redundant and probably
underwent competition processes.

@@

Validating the IDH, communities trajectories confirmed the diversity
increase after disturbance through an enhanced growth of otherwise less
favored species. This was strong regarding the functional diversity but
more blurry regarding their taxonomic structure.

The taxonomic richness was weakly or negatively impacted by intense
disturbance, as observed on several post logging surveys
\citep{Cannon1998, Baraloto2012a}, while it substantially increased
after low intensity disturbance. Disturbance also enhanced taxonomic
evenness, but the disturbance intensity was only weakly and lately
correlated to communities eveness (\emph{i.e.} Simpson diversity).
Contrastingly, disturbance intensity consistently predicted the
significant increase of communities functional diversity for 30 years.
Disturbance entailed specific turnover within communities, either among
pre-disturbance survivors or among newly recruited trees. As the
composition of old-growth survivors proved to mirror initial communities
\citep{Herault2018}, disturbance likely impacted trees recruitment
through the enhanced growth and survival of previously infrequent
species and functional types. Consistently, disturbance resulted in
increasing taxonomic dissimilarity compared to pre-disturbance
communities and significant functional shifts towards
resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. Disturbance then caused a
reorganization of the typical high dominance structure of hyperdiverse
mature forests after disturbance, benefiting to pioneers and light
demanding species. Likely, the changes in abiotic environment and
competitive pressure favored pioneers which outcompete other species in
non limiting resources but are excluded in mature forests through the
cmpetition with long-lived, more resistant and shade bearers. Therefore,
consistently with the IDH the trajectories after disturbance relied on
the environmental niches made available by disturbance filled by species
from a restrictied functional range. Recruited species therefore
differed from pre-disturbance ones, constituting a community all the
more diversified that the disturbance was intense \citep{Molino2001}.

@@

Both the lag between taxonomic and functional recovery and the middling
consistency of the IDH regarding communities taxonomic diversity made
central the role of functional redundancy. Functional redundancy,
determinant of communities resilience
\citep{Trenbath1999, Elmqvist2003, Diaz2005}, seemed not to have
recovered 30 years after disturbance.

Communities with high functional redundancy are expected to be able to
lose species with a limited impact on their functional structure
\citep{Carmona2016}. Functional redundancy therefore mitigates the
impacts of disturbance, but we expect the redundancy itself to be
reduced or re-organized within the functional space. Although the
overall functional redundancy did not follow consistent trajectories
after disturbance, the redundancy restricted to the functional space of
the initial community clearly followed humped shaped trajectories. This
restricted redundancy decreased according to the disturbance intensity
before returning towards the initial value but had not recovered for any
disturbed community after 30 years. The recovery of the redundancy
within the initial functional space explained the hampered recovery of
infrequent species specific to old-growth forests that would be
functionally redundant and therefore underwent competitive pressures
limiting species functional similarity \citep{Mayfield2010}. The
unachieved humped shaped redundancy trajectory meant a decrease of the
resilience of pre-disturbance communities and an increase in the
resilience of disturbance-specific communities (corresponding to lower
richness and higher dominance of pioneer species). After disturbance the
chances were higher to have long lasting or self-maintained
compositional changes in favor of disturbance resistant species, lianas
or epiphytes \citep{Haddad2008, Burslem2000, Martin2013}, which highly
questioned the completeness of the resilience \citep{Chazdon2003a}.
Specificaly, species contingent to undisturbed forests might be
impaired, threatening their maintenance, the risk to loose cornerstone
species and trigger unexpected ecological consequences would be higher
\citep{Jones1994, Diaz2005, Gardner2007}.

@@

our study defined the decoupled functional and taxonomic trajectories of
tropical rainforests after disturbance. These trajectories shed light on
their rapid functional recovery but slower and more variable taxonomic
dynamics. Consistently with the IDH, functional trajectories were
constrained by determined processes favoring pioneers and
light-demanding species after disturbance. The taxonomic dynamics
however proved more stochastic due to the recovery of infrequent,
functionally redundant species hampered by limiting similarity
processes. Although communities functioning rapidly recovered, their
taxonomic structure and functional redundancy remained altered for
several decades. The trajectories therefore demonstrated a decades-long
resilience of tropical forests in the face of quite intense disturbance
\citep{Gourlet-Fleury2005}. Within the mechanisms of this resilience the
recruitment processes proved central, in accordance with the IDH, and
would deserved closer focus to clarify their determinisms and ecological
rules. The disturbance range however stayed within the spectrum of
selective logging, with a forest cover remaining all along the
experiment, and the response mechanisms would probably be much different
after harder disturbance.

>>>>

Both communities taxonomic and functional diversity and composition
proved resilient, following similar humped shaped trajectories returning
towards initial values. The resilience of communities functional
structure, that is the most direct link between biodiversity and
ecosystem functioning \citep{Diaz2005}, meant a consistent recovery of
ecosystem processes in the long term \citep{Guariguata2001}. The
resilience of communities taxonomy meant the maintenance of their
initial differences in composition and diversity, suggesting the
existence of multiple stable equilibria as assumed for highly diverse
and productive ecosystems \citep{Chase2003}, and the dependency of
recovery trajectories on the initial composition to be recovered that
would the the asymptot of the humped shaped dynamics
\citep{Hubbell1999, Molino2001, Anderson2007, Baraloto2012a}.

Although both communities taxonomic and functional characteristics
proved resilient and followed similar humped shaped trajectories, the
taxonomic recovery systematically lagged behind the functional one. The
decoupling between functional and taxonomic dynamics has already been
observed for grasslands \citep{Tilman1997, Mouillot2011} and more
recently for tropical forests \citep{Lohbeck2015, Guariguata2001}.
According to the ``vegetation quantity effect'' \citep{Grime1998}
functional trajectories rely on the pool of dominant species, which
diversity and evenness increased after disturbance but rapidly recovered
their initial functional structure. Although the pool of dominant
species recovered quickly, communities evenness remained high so
infrequent species still missed to the taxonomic recovery. Taxonomic
recovery was then all the more longer that unrecovered species would be
functionally redundant and probably underwent selective competition
processes.

@@

In accordance with the IDH the trajectories confirmed the increase of
communities diversity after disturbance with the favored growth of
otherwise less favored species. However, although it was clear for the
functional diversity this tendency was more blurry regarding communities
taxonomic structure.

The taxonomic richness was enhanced after low disturbance but it was
weakly or negatively impacted by intense disturbance, as observed on
several post logging surveys \citep{Cannon1998, Baraloto2012a}. The
taxonomic evenness also increased after disturbance, but the correlation
between the disturbance intensity and the diversity increase remained
weak and valid only after 20 years (\emph{i.e.} Simpson diversity).

Contrastingly the functional diversity increase after disturbance was
determined by the disturbance intensity all along the 30 years
monitored. We already demonstrated that disturbance primarily entailed
turnover of specific within communities, either among pre-disturbance
survivors or among newly recruited trees. It wasdemonstrated that the
composition of old-growth survivors proved to mirror initial communities
\citep{Herault2018}, so disturbance rather impacted trees recruitment
with the enhanced growth and survival of previously infrequent species
and functional types. Consistently disturbance resulted in an increase
of the taxonomic dissimilarity between communities and their
pre-disturbance composition, and in significant functional shifts
towards resource-acquisitive strategies (sharp increase in the SLA, leaf
thickness and bark thickness and decrease in wood density, leaf
toughness and maximum height)
\citep{Westoby1998, Wright2004, Reich2014}. Disturbance entailed a
reorganization of the typical high dominance structure of hyperdiverse
mature forests benefiting to pioneers and light demanding species: the
changes in abiotic environment and competitive pressure favored pioneers
which outcompete other species in non limiting resources but were
excluded in mature forests by long-lived, more resistant and shade
tolerant species. Corroborating the IDH, the trajectories after
disturbance relied on the environmental niches made available and filled
by species from a restrictied functional range. Recruited species
therefore differed from pre-disturbance ones, constituting a community
all the more diverse that the disturbance was intense
\citep{Molino2001}.

@@

Both the lag between taxonomic and functional recovery and the middling
consistency of the IDH regarding communities taxonomic diversity were
explained by the incomplete recovery of communities functional
redundancy, which besides implies that communities resilience remained
altered 30 years after disturbance
\citep{Trenbath1999, Elmqvist2003, Diaz2005}.

Functional redundancy mitigates the loss of species on communities
functional structure \citep{Carmona2016}, and therefore reduces the
impact on disturbance. The functional redundancy itself however might be
reduced or re-organized within the functional space and alter
communities resilience. Although the functional redundancy considered in
the whole commnities functional space did not follow consistent
trajectories after disturbance, the redundancy restricted to the
functional space of the initial community clearly followed humped shaped
trajectories. This restricted redundancy first decreased, all the more
so that the disturbance was intense, and then returned towards the
initial value but this had not recovered for any disturbed community
after 30 years. After disturbance communities then differently occupied
the functional space, as it was suggested by the shifts observed for the
functional diversity and traits trajectories, before recovering the
specific highly redundant initial functional space. This increasing
redundancy along the trajectory explained the hampered recovery of
infrequent species specific to old-growth forests, as they were likely
functionally redundant and therefore underwent competitive pressures
limiting species functional similarity \citep{Mayfield2010}. The
unachieved recovery of the functional redundancy meant a decreased
resilience of pre-disturbance communities and an increased of
disturbance-specific communities (corresponding to lower richness and
higher dominance of pioneer species). After disturbance the chances were
therefore higher to have long lasting or self-maintained compositional
changes in favor of disturbance resistant species, lianas or epiphytes
\citep{Haddad2008, Burslem2000, Martin2013}. The recovery was not
complete until the functional redundancy and resilience of the initial
communities remained altered. Specificaly, such incomplete recovery
would threaten the maintenance of species contingent to undisturbed
forests and increase the risk of cornerstone species loss and unexpected
ecological consequences
\citep{Jones1994, Chazdon2003a, Diaz2005, Gardner2007}.

@@

Our study defined the decoupled functional and taxonomic trajectories of
tropical rainforests after disturbancethat demonstrated a rapid
functional recovery but a slower and more variable taxonomic one.
Consistently with the IDH, functional trajectories were driven by
deterministic processes favoring pioneers and light-demanding species
after disturbance. The following functional shifts entailed a
re-organization of communities functional redundancy in the functional
space that proved long to recover and explained the more variable and
longer taxonomic trajectories as shade tolerant species of old-growth
forests faced functional redundancy and limiting similarity processes.
The resilience of tropical forests then proved long in the face of quite
intense disturbance and did not preclude the settling of a persistent
disturbanc-specific community \citep{Gourlet-Fleury2005}. Within the
mechanisms of this resilience the recruitment processes proved central,
in accordance with the IDH, and would deserve closer focus to specify
their ecological drivers. The disturbance range however stayed within
the spectrum of selective logging, with a forest cover remaining all
along the experiment, and the response mechanisms would probably be much
different after harder disturbance.
